Download data

### 2023-12-14 scripts/download_dependent_variables.Rmd
downloaded misspelled taxa

```{r import and clean target species}
targets <- read.csv('../data/raw/2024sos_species.csv', na.strings = "") %>% 
  filter(!species %in% c('tridentata', 'millefolium', 'dioica', 'negundo')) %>%  # purshia,
  unite('taxon', genus:infraspecies, na.rm = T, sep = ' ') %>% # artemisia, and larrea !
  mutate(taxon = str_squish(taxon))

@@

## we will not model a few species which are widely abundant, hence do not require modelling, and would utilize enormous computational resources.

@@

gbif_taxon_keys <- target_taxa[294:364] %>% 
  name_backbone_checklist()  %>% 
  filter(!matchType == "NONE") %>% 
  pull(usageKey)

@@

occ_download_wait(dl_key)
occ_download_get( dl_key )

@@

for(i in seq_along(target_taxa)){
  print(target_taxa[i])
  dl <- BIEN_occurrence_species(target_taxa[i], collection.info = TRUE, observation.type = TRUE, new.world = TRUE)
  write.csv(dl, file.path('../data/raw/occurrence/bien', paste0(gsub(' ', '_', target_taxa[i]), '.csv')) )#, row.names = FALSE)
}
```

@@

@@

write.csv(idig, '../data/raw/occurrence/IDIG.csv')

@@

spp_richness <- st_read(dsn = p1, 
        layer = 'tblSpecRichDetail', quiet = T) %>% 
  select(SpeciesList, PrimaryKey) %>% 
  filter(SpeciesList %in% target_symbols) %>% 
  right_join(AIM_points, by = 'PrimaryKey') %>% 
  st_as_sf()  %>% 
  mutate(DateVisited = as.Date(DateVisited), .before = geometry) %>% 
  arrange(SpeciesList) %>% 
  right_join(., select(targets, taxon, symbol) %>% 
               distinct(), by = c('SpeciesList' = 'symbol' ))

st_write(spp_richness, '../data/raw/occurrence/AIM/AIM_records.shp', append = F)

@@

@@

```{r remove points outside analytical area}
library(terra)
domain <- rast(nrows = 1, ncols = 1) # create a big empty raster, you can go in through sf too. 
ext(domain) <- c( -125.5, -100, 27,  50) # set the extent
crs(domain) <- "EPSG:4326" # define the projection

@@

>>>>

```{r import and clean target species}
targets <- read.csv('../data/raw/2024sos_species.csv', na.strings = "") %>%

  ## we will not model a few species which are widely abundant, hence do not require modelling, and would utilize enormous computational resources.
  filter(!species %in% c('tridentata', 'millefolium', 'dioica', 'negundo')) %>%  # purshia,
  unite('taxon', genus:infraspecies, na.rm = T, sep = ' ') %>% # artemisia, and larrea !
  mutate(taxon = str_squish(taxon))

@@

## these species had initial errors, fix on Dec 14th, 
target_taxa <- c('Brodiaea elegans', 'Krameria bicolor', 'Lotus humistratus', 'Adenostoma fasciculatum')

@@

gbif_taxon_keys <- target_taxa %>% 
  name_backbone_checklist()  %>% 
  filter(!matchType == "NONE") %>% 
  pull(usageKey)

@@

occ_download_wait(dl_key)
occ_download_get(dl_key)

@@

for(i in seq_along(target_taxa)){
  print(target_taxa[i])
  dl <- BIEN_occurrence_species(target_taxa[i], collection.info = TRUE, observation.type = TRUE, new.world = TRUE)
  write.csv(dl, file.path('../data/raw/occurrence/bien', paste0(gsub(' ', '_', target_taxa[i]), '.csv')) )#, row.names = FALSE)
}

rm(i)
```

@@

```{r stitch together BIEN queries 2, eval = F}

p <- '../data/raw/occurrence/bien'
files <- file.path(p, list.files(p))
# files <- files[ grep("Adenostoma_fasciculatum.csv|Brodiaea_elegans.csv|Krameria_bicolor.csv|Lotus_humistratus.csv", files) ]

write.csv(bien_data, '../data/raw/occurrence/BIEN_records2.csv', row.names = FALSE)
```

@@

write.csv(idig, '../data/raw/occurrence/IDIG2.csv')
```

@@

target_symbols <- c('ADFA', 'ADFAF', 'ADFAO', 'BREL', 'BRELE', 'KRGR', 'LOHU2')
targets <- filter(targets, symbol %in% target_symbols) %>% 
  distinct()

spp_richness <- st_read(dsn = p1, 
        layer = 'tblSpecRichDetail', quiet = T) %>% 
  select(SpeciesList, PrimaryKey) %>% 
  # perform this a second time with new targets #
  filter(SpeciesList %in% target_symbols) %>% 
  right_join(AIM_points, by = 'PrimaryKey') %>% 
  st_as_sf()  %>% 
  mutate(DateVisited = as.Date(DateVisited), .before = geometry) %>% 
  arrange(SpeciesList) %>% 
  right_join(., select(targets, taxon, symbol) %>% 
               distinct(), by = c('SpeciesList' = 'symbol' ))

st_write(spp_richness, '../data/raw/occurrence/AIM/AIM_records2.shp', append = F)

@@

```{r Create consensus presence records 1}

@@

bien <- read.csv(file.path(base_p, 'BIEN_records2.csv')) %>% 
  st_as_sf(coords = c('longitude', 'latitude'), crs = 4326) %>% 
  mutate(date = as.Date(date_collected)) %>% 
  select(taxon = scrubbed_species_binomial, date)

idig <- read.csv(file.path(base_p, 'IDIG2.csv')) %>% 
  st_as_sf(coords = c('geopoint.lon', 'geopoint.lat'), crs = 4326) %>% 
  mutate(date = as.Date(datecollected)) %>% 
  select(taxon = scientificname, date)

aim <- st_read(file.path(base_p, 'AIM', 'AIM_records2.shp'), quiet = TRUE) %>% 
  mutate(date = as.Date(DatVstd)) %>% 
  select(taxon, date, PrmryKy) %>% 
  st_transform(4326)

gbif3 <- read.delim(file.path(base_p, 'gbif3', 'occurrence.txt'), na.strings = "") %>% 
  drop_na(decimalLongitude, decimalLatitude) %>% 
  filter(str_detect(decimalLatitude, '[A-Z]|[a-z]', negate = T)) %>% 
  st_as_sf(coords = c('decimalLongitude', 'decimalLatitude'), crs = 4326) %>% 
  mutate(date = as.Date(eventDate), .before = geometry, 
         coordinateUncertaintyInMeters = as.numeric(coordinateUncertaintyInMeters)) %>% 
  filter(coordinateUncertaintyInMeters <= 250) %>% 
  select(taxon = species, date)

records <- bind_rows(aim, bien, gbif3,  idig) %>% 
  mutate(taxon = str_to_sentence(taxon), 
         taxon = str_squish(taxon), 
         taxon = str_replace(taxon, 'Adenostoma fasiculatum', 'Adenostoma fasciculatum'), 
         taxon = str_replace(taxon, 'Brodiaeae elegans', 'Brodiaea elegans'), 
         taxon = str_replace(taxon, 'Lotus humistratus|Acmispon humistratus', 'Acmispon brachycarpus')
         ) %>% 
  arrange(taxon) # recs need be > 90 m from each other

rm(aim, bien, gbif3,  idig)

splicies <- split(records, records$taxon)
thinned <- lapply(splicies, ONEperCELL)
thinned <- bind_rows(thinned)
st_write(thinned, '../data/raw/occurrence/thinned/thinned_occurrences2.shp', append = F)
repeated_occurrences <- lapply(splicies, multiplesPERcell)
repeated_occurrences <- purrr::discard(repeated_occurrences, ~nrow(.) < 2)
repeated_occurrences <- bind_rows(repeated_occurrences)
st_write(repeated_occurrences, '../data/raw/occurrence/repeated/repeated_occurrences2.shp', append = F)

```{r remove points outside analytical area, eval = F}
library(terra)
domain <- rast(nrows = 1, ncols = 1) # create a big empty raster, you can go in through sf too. 
ext(domain) <- c( -125.5, -100, 27, 50) # set the extent
crs(domain) <- "EPSG:4326" # define the projection

@@

```{r remove points outside analytical area2, eval = F}
library(terra)
domain <- rast(nrows = 1, ncols = 1) # create a big empty raster, you can go in through sf too. 
ext(domain) <- c( -125.5, -100, 27, 50) # set the extent
crs(domain) <- "EPSG:4326" # define the projection

thinned <- st_read('../data/raw/occurrence/thinned/thinned_occurrences2.shp', quiet = TRUE)
thinned <- thinned[lengths(st_disjoint(thinned, domain)) ==  0, ]
st_write(thinned, '../data/raw/occurrence/thinned/thinned_occurrences2.shp', append = FALSE)

repeated_occurrences <- st_read('../data/raw/occurrence/repeated/repeated_occurrences2.shp', quiet = TRUE)
repeated_occurrences <- thinned[lengths(st_disjoint(repeated_occurrences, domain)) ==  0, ]
st_write(repeated_occurrences, '../data/raw/occurrence/repeated/repeated_occurrences2.shp', append = FALSE)
```
